---
format:
  html:
    embed-rmd: true
    output-ext: "htm"
---

# Example: Figures in R and Python

This chapter demonstrates how to organize multiple figures from R and Python in interactive tabs using Quarto.

:::: {.panel-tabset}

### R: Scatter Plot Example

```{r}
library(ggplot2)

# Generic example data
set.seed(123)
df <- data.frame(
  X = rnorm(100),
  Y = rnorm(100),
  Group = sample(c("A", "B"), 100, replace = TRUE)
)

ggplot(df, aes(X, Y, color=Group)) +
  geom_point(size=2) +
  labs(title="R Scatter Plot", x="X axis", y="Y axis", color="Group") +
  theme_minimal()
```

**Figure:** Example scatter plot with two groups.  
[Download PDF](report_figures/RScatterPlot.pdf)

---

### R: Histogram Example

```{r}
ggplot(df, aes(X, fill=Group)) +
  geom_histogram(bins=20, alpha=0.7, position="identity") +
  labs(title="R Histogram", x="X axis", y="Count", fill="Group") +
  theme_minimal()
```

**Figure:** Example histogram by group.  
[Download PDF](report_figures/RHistogram.pdf)

---

::::

## Python Figures

:::: {.panel-tabset}

### Python: Sine Curve Example

```{python}
import matplotlib.pyplot as plt
import numpy as np

x = np.linspace(0, 2 * np.pi, 100)
y = np.sin(x)

plt.figure(figsize=(6, 4))
plt.plot(x, y, label='sin(x)', color='blue')
plt.title("Python Sine Curve")
plt.xlabel("x")
plt.ylabel("sin(x)")
plt.legend(title="Legend")
plt.show()
```

**Figure:** Example sine curve plot.  
[Download PDF](report_figures/PythonSineCurve.pdf)

---

### Python: Bar Plot Example

```{python}
import matplotlib.pyplot as plt

categories = ['A', 'B', 'C']
values = [10, 15, 7]

plt.figure(figsize=(6, 4))
plt.bar(categories, values, color=['red', 'green', 'blue'])
plt.title("Python Bar Plot")
plt.xlabel("Category")
plt.ylabel("Value")
plt.legend(['Values'], title="Legend")
plt.show()
```

**Figure:** Example bar plot for categories.  
[Download PDF](report_figures/PythonBarPlot.pdf)

::::